---
title: "EDA"
author: "aigerim"
date: "27 09 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(scipen = 999)
library(data.table)
library(tidyverse)
library(GGally)
library(skimr)
library(plotly)
library(memoise)
library(scales)
library(tsbox)
library(treemap)
library(fuzzyjoin)
library(ggthemes)
library(maps)
library(forecast)
theme_set(theme_light())
```

```{r}
mem_parse_date <- memoise(parsedate::parse_date)
```



```{r}
hotels_raw <- fread('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv') 
skim(hotels_raw)
#colSums(is.na(hotels_raw)) # not many NA's in the traditional form 
#(hotels_raw$company %>% table())/nrow(hotels_raw) # but a lot in the form of "NULL"
```

```{r}
hotels_cleaned <- hotels_raw %>%
  dplyr::select(-company) %>% #90% + of values are na's
  mutate(arrival_date = paste(arrival_date_day_of_month, arrival_date_month, arrival_date_year) %>% mem_parse_date()) %>% #date of arrival
#Mutate adds new variables and preserves existing
  mutate(total_number_of_people = adults + children + babies) %>%
  mutate(children_and_babies = children + babies) %>%
  mutate(country = if_else(country == 'CN', 'CAN', country)) %>%
  mutate(total_nights = stays_in_weekend_nights + stays_in_week_nights) %>%
  mutate(total_nights_cat = total_nights %>% as.factor() %>% fct_lump(., 5)) %>%
  mutate(required_car_parking_spaces_binary = if_else(required_car_parking_spaces > 0, T,F)) %>%
  dplyr::filter(adr %between% c(0,600)) %>%
  na.omit()

hotels_cleaned_not_cancelled <- hotels_cleaned[is_canceled == 0]
```


```{r}
plot_people_total <- hotels_cleaned_not_cancelled  %>%
  group_by(arrival_date, hotel) %>% 
  summarize(people_per_day = sum(total_number_of_people)) %>%
  ungroup() %>%
  ggplot(aes(x = arrival_date, y = people_per_day, color = hotel)) + 
  geom_line() + 
  facet_wrap(~hotel)

ggp_build <- plotly_build(plot_people_total)
ggp_build
```

```{r}
# August is the most occupied (busiest) month with 11.49% bookings and January is the most unoccupied month with 5.48% bookings.
hotels_cleaned_not_cancelled  %>%
  count(arrival_date_month) %>%
  rename(how_many_guests_arrived = n) %>%
  mutate(per_of_total = (100*how_many_guests_arrived / nrow(hotels_cleaned_not_cancelled)) %>% round(., 2) %>% paste(., '%'))
  
```

```{r}
plot_children <- hotels_cleaned_not_cancelled  %>%
  group_by(arrival_date, hotel) %>% 
  summarize(children_per_day = sum(children)) %>%
  ungroup() %>%
  ggplot(aes(x = arrival_date, y = children_per_day, color = hotel)) + 
  geom_line()

ggp_build <- plotly_build(plot_children)
ggp_build
```


```{r}
# as expected, the number of children and babies reaches its peak during summer months (July and August) and drops quickly before September

children_and_babies_ts <- hotels_cleaned_not_cancelled  %>%
  group_by(arrival_date) %>% 
  summarize(children_per_day = sum(children_and_babies)) %>% 
  ungroup() %>% 
  mutate(arrival_date = as.Date(arrival_date)) %>% 
  ts_ts() 

ggplotly(
  children_and_babies_ts %>% 
  mstl() %>%
  autoplot()
)

```

```{r}
#most guests (92%) of the two hotels do not have children and babies, while some have 1 or 2. Also there were rows with 9 and 10 children - probably groups 
hotels_cleaned_not_cancelled %>% 
  count(children_and_babies) %>% 
  mutate(percent = round(n / nrow(hotels_cleaned_not_cancelled), 2))

```

```{r}
#50% of the guests book their hotel approximately 45 days before the trip or less 
hotels_cleaned_not_cancelled$lead_time %>% quantile(c(0.25, 0.5, 0.75))

hotels_cleaned_not_cancelled %>% 
  ggplot(aes(x = lead_time)) + 
  geom_histogram() + 
  geom_vline(xintercept = median(hotels_cleaned_not_cancelled$lead_time), color = 'red') + 
  scale_x_log10()
```

```{r}
# Out of the meals, BB (Bed & Breakfast) is the most ordered meal which is around 77.2%, followed by HB(Half Board), SC(no meal package), Undefined and FB (Full Board).
hotels_cleaned_not_cancelled %>% 
  count(meal) %>% 
  mutate(percent = round(n / nrow(hotels_cleaned_not_cancelled), 2)) %>%
  mutate(meal = fct_reorder(meal,n)) %>%
  ggplot(aes(x = meal, y = n, fill = meal, label = percent)) + 
  geom_col() + 
  geom_label()
```
```{r}
# Around 40% of all bookings were booked from 
# Portugal followed by Great Britain(10%) & France(8%)

world_map <- map_data('world') %>%
  as.data.table() %>%
  fuzzyjoin::regex_left_join(maps::iso3166 %>% select(mapname, country_code = a3), by = c('region' = 'mapname')) %>%
  dplyr::filter(region != 'Antarctica')

world_map_plot <- world_map %>%
  left_join(hotels_cleaned %>% count(country), by = c('country_code' = 'country')) %>%
  ggplot(aes(long, lat, group = group, fill = n, text = region)) + 
  geom_polygon() +
  scale_fill_gradient2(low = "yellow", mid = 'blue', high = 'red', midpoint = 1e4) + 
  theme_map() 

ggplotly(world_map_plot)
```
```{r}
#the statistics with actual visitors is very similar

world_map <- map_data('world') %>%
  as.data.table() %>%
  fuzzyjoin::regex_left_join(maps::iso3166 %>% select(mapname, country_code = a3), by = c('region' = 'mapname')) %>%
  dplyr::filter(region != 'Antarctica')

world_map_plot <- world_map %>%
  left_join(hotels_cleaned_not_cancelled %>% count(country), by = c('country_code' = 'country')) %>%
  ggplot(aes(long, lat, group = group, fill = n, text = region)) + 
  geom_polygon() +
  scale_fill_gradient2(low = "yellow", mid = 'blue', high = 'red', midpoint = 1e4) + 
  theme_map() 

ggplotly(world_map_plot)
```


```{r}
#Around 47% of bookings are made via Online Travel Agents, almost 20% of bookings are made via Offline Travel Agents and less than 20% are Direct bookings without any other agents.
hotels_cleaned %>%
  group_by(market_segment) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(market_segment = fct_reorder(market_segment,n)) %>%
  ggplot(aes(x = market_segment, y = n)) + 
  geom_col()

```

```{r}
# Most of the customers belong to the Transient type 
hotels_cleaned_not_cancelled %>%
  group_by(customer_type) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(customer_type = fct_reorder(customer_type,n)) %>%
  ggplot(aes(x = customer_type, y = n)) + 
  geom_col()

```

```{r}
#Most people stay only for 1-3 nights
hotels_cleaned_not_cancelled %>% 
  count(total_nights_cat) %>%
  ggplot(aes(x = total_nights_cat, y = n)) + 
  geom_col() + 
  coord_flip()
```

```{r}
#most people do not have any special requests, while some have 1 or 2
hotels_cleaned_not_cancelled %>% 
  mutate(total_of_special_requests = total_of_special_requests %>% as.factor() %>% fct_lump(.,3)) %>%
  count(total_of_special_requests) %>%
  ggplot(aes(x = total_of_special_requests, y = n )) + 
  geom_col() + 
  coord_flip()
```
```{r}
#37% of the bookings are canceled
(hotels_cleaned$is_canceled %>% table() %>% c())/nrow(hotels_cleaned)
```

```{r}
#almost everybody in the dataset is a new guest 
(hotels_cleaned_not_cancelled$is_repeated_guest  %>% table() %>% c())/nrow(hotels_cleaned_not_cancelled)
```
```{r}
#most people do not require parking spaces
hotels_cleaned_not_cancelled %>% count(required_car_parking_spaces_binary)
```

```{r}
#types 'A' and 'D' are by far the most popular (the hotels that provided data for this dataset didn't specify the differences between room types for anonymity resons)
hotels_cleaned_not_cancelled %>% 
  count(reserved_room_type) %>%
  treemap(.,
            index="reserved_room_type",
            vSize="n",
            type="index",
            title = 'Reserved Room types' 
            )
```

```{r}
#Average Daily Rate as defined by dividing the sum of all lodging transactions by the total number of staying nights
#most people are near the value of 100 with a log of outliers including a negative one (should be excluded) and a very positive one
#so i included a filtering step to the processing of the data 
hotels_raw %>%
  ggplot(aes(y = adr)) + 
  geom_boxplot() + 
  scale_y_log10() + 
  coord_flip()

hotels_raw %>% arrange(adr) 
hotels_raw %>% arrange(desc(adr)) 
```

```{r}
# adr is positively  and stat.sign correlated with the number of children + babies 
# adr is positively  and stat.sign correlated  with the number of total requests
# the number of total requests is positively  and stat.sign correlated with the number of children + babies

hotels_cleaned_not_cancelled %>%
  select(adr, children_and_babies, required_car_parking_spaces_binary, total_of_special_requests) %>% 
  ggpairs(aes(fill = children_and_babies))

```

After concluding the EDA, we could formulate two goals that might be useful for hotels and other related businesses
1) Be able to predict what rooms are being booked by different people in order to better coordinate marketing (what customers do we target) and logistical/construction strategies (how many rooms of different types we need to have). For example, the top management decided to increase the focus on the business audience. A working classificator for reserved rooms type would provide an opportinity to estimate the expected distribution of reserved rooms for the given distribution of customes (simulation) 
2) Be able to predict whether or not the customer is going to cancel the booking in order to predict the loading of the hotel slightly better?







